---
title: 'Part 1: Simulation Exercise'
author: "David Augusto"
date: "25/10/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


#### Part 1: Simulation Exercise Instructions
 
In this project you will investigate the *exponential distribution* in R and compare it with the Central Limit Theorem.


The exponential distribution can be simulated in R with **rexp(n, lambda)** where:

1. **lambda** is the rate parameter. 

2. The *mean* of exponential distribution is **1/lambda**. 

3. the *standard deviation* is also **1/lambda**. 

Set **lambda = 0.2** for all of the simulations.




*You will investigate the distribution of averages of 40 exponentials. *


Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. 

You should:

1. Show the **sample mean** and *compare it* to the **theoretical mean** of the distribution.

2. Show the **variance** and *compare it* to the **theoretical variance** of the distribution.

3. **Show that the distribution is approximately normal.**


 ======================================================================================================================

#### Exponential distribution:


The exponential distribution is defined by its cumulative distribution function

$$ F(x) = 1-e^{-\lambda x} $$

The R function `rexp` generates random variables with an exponential distribution.






>Question 1 : Show the sample mean and compare it to the theoretical mean distribution

```{r}

set.seed(777)
n <- 40
Lambda <- 0.2
Simulations <- 1000

```


>Theoretical values:


1. Theoretical Mean is equal to: `r 1/Lambda`

2. Theoretical Standard Deviation is equal to: `r 1/Lambda* (1/sqrt(n))`

3. Theoretical Variance is equal to: `r (1/Lambda* (1/sqrt(n)))^2`

```{r}

SampleMean <- 0

#For,this will create our n means from the rexp funtion.

for(i in 1:Simulations) {
  SampleMean <- c(SampleMean, mean(rexp(n, Lambda)))
  
}

mean(SampleMean)
```

The sample mean is very close to the theoretical mean distribution.
The difference is = `r 5-4.969789`


>Question 2: Show the sample variance and compare it to the thoretical variance of the distribtution.

```{r}
Variance <- var(SampleMean)

Variance
```


The sample variance is very close to the theoretical variance.

The difference is = `r .625-0.6384832`




**Show that the distribution is aproximately normal**


This can be shown with 2 graph tools:

1. Histogram of the n means

2. The normal probability plot 

======================================================================================================================

First we examine the Histogram, 

```{r}

hist(SampleMean, breaks = n, prob = T, col = "green", xlab = "Means")
x <- seq(min(SampleMean), max(SampleMean), length = 100)
lines(x, dnorm(x, mean = 1/Lambda, sd = (1/Lambda/sqrt(n))), pch = 25, col = "Red")


```

We conclude that the histogram appears to be bell shaped, and the mean is very close to the theoretical mean.


Then, for our Normal probability plot:


```{r}
qqnorm(SampleMean)
qqline(SampleMean, col = "Red")
```

The data should form an approximate straight line, in this case it appears almost linear, we can assume the data set **is approximately normally distributed**


